---
title: "p8105_hw2_mbc2178"
author: "Melvin Coleman"
date: "2022-10-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's load in the libraries needed to perform data manipulation and cleaning for this assignment.
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```

## Problem 1

Below we import the csv file containing NYC transit data `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. We first import our dataset to R and perform data cleaning and manipulations to fit the assignment. We keep the variables 
`station`, `name`, `station latitude/longitude`, `routes`, `entry`, `vending`, `entrance type`, and `ada`. 
In addition, we convert the variable `entry` from character to a logical variable. 

```{r import}
nyc_transit_df = 
    read_csv(
      file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
      col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
     janitor::clean_names() %>%
    select(
      line, station_name, station_latitude, station_longitude, starts_with("route"),
      entry, vending,  exit_only, vending, entrance_type, ada) %>%
    mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

After completing the cleaning procedure above, the data still doesn't look tidy.
There are `r nrow(nyc_transit_df)`and `r ncol(nyc_transit_df)`. The variables `route8 - route11`
were changed to character variables and the variable `entry` was changed to a logical variable. 

The variables in our dataset include `line`, `station_name`, `station_latitude`, `station_longitude`,
`route1 - route11`,`entry`, `vending`, `exit_only`, `vending`, `entrance_type` and `ada`. 

We will need to perform additional manipulations on variables `route1-route11` to ensure that there is one variable that
contains all of the subway routes. 

The code chunk below selects the station names and line numbers to determine the number of distinct stations. We 
use the `select` function to select `station_name` and `line` and use the `distinct` function to only select a unique 
station per row in our data frame. 

```{r}
nyc_transit_df %>% 
  select(station_name, line) %>% 
  distinct
```
There are 465 distinct stations in the `nyc_transit_df`(the number of rows in the table are 465).

Next, we determine the number of stations that are ADA compliant in this dataframe. We first use the 
`filter` function to filter out only stations that are ADA compliant and select the unique 
station and line in the subway system per row. 


```{r}
nyc_transit_df %>% 
  filter(ada == "TRUE") %>% 
  select(station_name, line) %>% 
  distinct()
```
 There are 84 stations that are ADA compliant in the `nyc_transit_df` data frame (the number of rows in this tibble is 84).

Now we determine the proportion of station entrances/ exits without vending allow entrance. We exclude stations that 
do that allow vending by using the `filter` function. Then we use the `pull` function on `entry` to find the mean of 
this logical variable. This mean will tell us the proportion of station entrances/exists without vending allow entrance. 

```{r}
nyc_transit_df %>%
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

There are 37.7% of station entrances/exits without vending allow entrance (mean from code chunk output above). 

Last, we now reformat our data so that the route number and route name are distinct variables. We use the `pivot_longer` to 
transform the variables to rows. Then we filter to select only the A train and use the `distinct function` to select 
unique stations that serve the A train and the number that are ADA compliant. 

```{r}
nyc_transit_df %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>%
select(station_name,line)%>%
distinct()

nyc_transit_df %>%
   pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route =="A", ada =="TRUE") %>%
select(station_name,line) %>%
distinct()
```
There are 60 stations that serve the A train and 17 stations re ADA compliant in the `nyc_transit_df` dataframe.

## Problem 2 

Let's load the `readxl` library to import the excel sheet containing data for Mr. Trash Wheel dataset. 
We will clean this dataset by ommitting non-data entries using arguments in the `read_excel` 
function. Then we will round variable `sports_balls` to the nearest integer using `as.integer`.
We will perform a similar process on the excel sheet containing data for Professor Trash Wheel. 

After data cleaning, we will combine the two datasets to produce a single tidy dataset. 
We will create a unique variable called `loc_trash` to keep track of which Trash Wheel is which. 


```{r}
library(readxl)
trash_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", 
             range = "A2:N534", na = "", col_names = TRUE, col_types = NULL) %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls, digits=0))) %>% 
  mutate( loc_trash = "Mr. Trash Wheel")
```

```{r}
proftrsh_df =
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
             sheet = "Professor Trash Wheel", range = "A2:N117" , na = "",
             col_names = TRUE, col_types = NULL) %>% 
janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls, digits=0))) %>% 
  mutate(loc_trash = "Professor Trash Wheel")
```
 

```{r join}

```




